---
title: "Leaflet Tutorial"
output:
  html_document:  
    toc: true  
    toc_float: true 
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(httr)
library(jsonlite)
library(dplyr)
library(tidyr)
library(extrafont)
library(leaflet)
library(kableExtra)

#access url
base2 <- "http://api.letsride.co.uk/public/v1/rides"
openactive <- GET(base2) 

#from unicode to text
this.raw.content <- rawToChar(openactive$content)

#parse to JSON
this.content <- fromJSON(this.raw.content)

#extract geo information 
newdata <- this.content$items
newdata1 <- newdata$data
newdata2 <- newdata1$location
newdata3 <- newdata2$geo

#bind dataframes
newdata4 <- cbind(newdata1, newdata3)


#rename Gender categories
newdata4$genderRestriction[newdata4$genderRestriction == "https://openactive.io/FemaleOnly"] <- "Women-only Session"
newdata4$genderRestriction[is.na(newdata4$genderRestriction)] <- "Mixed-Session"

mses <- subset(newdata4, genderRestriction == "Mixed-Session") 
wses <- subset(newdata4, genderRestriction == "Women-only Session")

```

# About Leaflet
The `Leaflet` package creates a map widget; you can add layers such as tiles, markers, and polygons to build up the map. There is already a fantastic guide for using [Leaflet](https://rstudio.github.io/leaflet/), but here is a quick overview on how to create a simple map, using data from British Cycling via the [Open Active API]("http://api.letsride.co.uk/public/v1/rides"). 

# Load the Library

All we need to create the map is a dataset with longitude and latitude (which we have from the Open Active API) and the `Leaflet` package.

```{r library, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(leaflet)
```

# Get the data

Here's a table of my British Cycling session information:

```{r table, echo=FALSE, fig.align='center'}
kable(newdata4[1,c(27,35,36)]) %>%
  kable_styling(bootstrap_options = "bordered",
                full_width = FALSE) 
```


# Create your map

Below is a map widget with Open Street Map tiles (the default) and circle markers of different colours to differentiate between mixed sessions, and women-only sessions:

```{r pressure, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.align='center'}
# use leaflet to map
leaflet() %>% addTiles() %>%
  addCircleMarkers(data= mses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction), color = "#43A8B6")  %>%
  addCircleMarkers(data= wses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction),  color = "#5D4086")
```

It is also possible to cluster these markers by using the `clusterOptions` argument:

```{r leaf, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.align='center'}
leaflet() %>% addTiles() %>%
  addCircleMarkers(data= mses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction), color = "#43A8B6", clusterOptions = markerClusterOptions())  %>%
  addCircleMarkers(data= wses, ~longitude, ~latitude, popup = ~as.character(genderRestriction), label = ~as.character(genderRestriction),  color = "#5D4086", clusterOptions = markerClusterOptions())

```

# Add layers and use custom markers 

There is an add-on package to the `leaflet` package = `leaflet.esri` - which enables a user to add Esri services to a `leaflet` map. The below code provides an example of adding feature layers from the Esri service, and adding a control for both tiles and layers.

It is also possible to create custom markers through the use of the [Font Awesome](https://fontawesome.com/) icon set.

```{r esri, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.align='center'}
library(leaflet.esri) # install 

woman <- makeAwesomeIcon(icon = "fa-female", library = "fa", markerColor = "blue") 
both <- makeAwesomeIcon(icon = "fa-bicycle", library = "fa", markerColor = "purple") 
#create marker

m <- leaflet()%>%
  addProviderTiles(providers$Esri.WorldGrayCanvas, group = "Esri Gray (default)") %>%
  addTiles(group = "OSM") %>%
  addProviderTiles(providers$OpenStreetMap.BlackAndWhite, group = "OSM Black and White") %>%
  addAwesomeMarkers(data= wses, ~longitude, ~latitude, label = ~as.character(genderRestriction), icon = woman) %>%
  addAwesomeMarkers(data= mses, ~longitude, ~latitude, label = ~as.character(genderRestriction), icon = both) %>%
  addEsriFeatureLayer(url = "https://services3.arcgis.com/ivmBBrHfQfDnDf8Q/ArcGIS/rest/services/Indices_of_Multiple_Deprivation_(IMD)_2019/FeatureServer/0",
                     useServiceSymbology = TRUE,
                     labelProperty = paste("LSOA01NM"), 
                     group = "IMD 2019",
                     labelOptions = labelOptions(textsize = "12px"),
                     popupProperty = JS("function(feature) { return L.Util.template(\"<h3>IMD Decile: {IMDDec0}</h3><hr /><p>This LSOA is {LSOA01NM} and the IMD Rank is {IMDRank0}. The total population is {TotPop} (excluding pensioners)\", feature.properties);}")) %>%
addEsriFeatureLayer(url = "https://services.arcgis.com/WQ9KVmV6xGGMnCiQ/ArcGIS/rest/services/UK_Health/FeatureServer/3",
                    useServiceSymbology = TRUE,
                    labelProperty = paste("AREA_NAME"), 
                    group = "Health UK 2011 Census",
                    labelOptions = labelOptions(textsize = "12px"),
                    popupProperty = JS("function(feature) { return L.Util.template(\"<h3>All categories: Long-term health problem or disability: {KS301EW0001}</h3><hr /><p>Day to day activities limited alot: {KS301EW0002} </br> Day-to-day activities limited a little: {KS301EW0003} </br> Day-to-day activities not limited: {KS301EW0004} \", feature.properties);}"))
  
#create controls 

m %>%
  addLayersControl(
    baseGroups = c("Esri Gray (default)", "OSM","OSM Black and White"),
    overlayGroups = c("IMD 2019", "Health UK 2011 Census"),
    options = layersControlOptions(collapsed = FALSE)
  )

```

