---
title: Small Grants in Focus
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: flatly
    highlight: tango
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(class)
library(gmodels)
library(rlang)
library(caret)
library(extrafont)
library(reshape2)
library(leaflet)
library(ggplot2)
library(dplyr)
library(extrafont)
library(leaflet)
library(scales)
library(knitr)
library(tidyverse)
library(forcats)
library(wordcloud2)
library(DT)
library(readxl)
library(kableExtra)

#### read in GMS data
setwd("C:/Users/jenniferb/OneDrive - Sport England/GitHub/Small Grants")
sg_df <- read.csv('SG_all.csv', header = TRUE, na.strings = c("","N/A"))

topic_tbl <- read.csv('12Month_top.csv') %>%
  rename("Topic" = Topic.id, "Terms" = Concatenate.Term.)

text_df <- read.csv('12Month_Prog.csv', header = TRUE)

gov <- read.csv('Gov.csv', header = TRUE, na.strings = c("","N/A"))

sentiment_df <- read.csv('Sentiment.csv', header = TRUE)%>%
  select("Document", "Sentiment.Prediction", "all.Words") %>%
  rename("Sentiment" = Sentiment.Prediction, `Number of Words` = "all.Words")

#### M&E data
me_df <- read_xlsx('MandE.xlsx', na = "NA")

##mean data
me_df_mean  <- me_df %>%
  select( "YEAR 1 PARTICIPANTS TARGET", "12 MONTH PARTICIPANTS") %>%
  mutate(`NUMBER OVER/UNDER` = `12 MONTH PARTICIPANTS` - `YEAR 1 PARTICIPANTS TARGET`)

## for women
me_df_women <- read_xlsx('MandE.xlsx', na = "NA")%>%
  select("MALE TARGET NUMBER", "MALE TARGET PERCENTAGE", "FEMALE TARGET NUMBER", "FEMALE TARGET PERCENTAGE", "12 MONTH MALE NUMBER", "12 MONTH FEMALE NUMBER", "12 MONTH MALE PERCENTAGE", "12 MONTH FEMALE PERCENTAGE",  "prediction" ) 



# colours for charts
se_colour<- c("#E79AB5", "#B8345C", "#8CCDD6", "#43A8B6", "#00818F", "#C1DC9A", "#7ABB40", "#5D4086")


# transformation for headline figs
sp_vec <- sg_df %>% count(Sport, sort = TRUE)
sp_per <- sp_vec[1,2] / sum(sp_vec$n) * 100
sp_per <- unlist(sp_per)
sp_org <- sg_df %>% count(ORGANISATION_NAME, sort = TRUE) 
sp_org$Times_funded <- ifelse(sp_org$n > 1, "More than once", "Once")
sp_la <- sg_df%>% count(LOCAL_AUTHORITY, sort = TRUE)
sp_fo <- sg_df%>% count(Focus, sort = TRUE)
sp_la_imd <- sg_df[order(sg_df$IMD_RANK),]
sent_tab <- table(sentiment_df$Sentiment)
gov_ord <- gov[order(gov$AWARD, decreasing = TRUE),]

## for sport chart
sg_sport <- sg_df %>%
  select(AWARD, Sport) %>%
  group_by(Sport) %>%
  summarise(AWARD = sum(AWARD))

### create count
count_of_sp <- sg_sport %>%
count(Sport)

count_of_sp <- sum(count_of_sp$n /2)

## for org chart
sg_org <- sg_df %>%
  select(AWARD, Org_Class) %>%
  group_by(Org_Class) %>%
  summarise(AWARD = sum(AWARD))

## for M&E
tab_me <- table(me_df$prediction)



```

# Introduction

Sport England is a funding body. Sending money out the door to support the sector achieve our outcomes is the lifeblood of what we do as an organisation. 
We recieve numerous applications from a diverse cast of characters, some of which fit the mould of "traditional" operators within this sector, others which may come to us from a slightly different backgrount. All of these applicants go through rigorous assessments carried out by our colleagues in the Investment Management Team. 

<br>

# Headline figures  {.tabset .tabset-fade .tabset-pills} 
These figures are for projects (funded by the Small Grants fund) that have started during our current strategy period (from January 2016 - November 2019):

* Number of awards made: `r nrow(sg_df)`
* Total amount awarded: `r paste0("£", formatC(sum(sg_df$AWARD), format="f", digits=2, big.mark=","))`
* Average amount awarded: `r paste0("£", formatC(mean(sg_df$AWARD), format="f", digits=2, big.mark=","))`
* Average total project cost: `r paste0("£", formatC(mean(sg_df$Total_Project_Cost), format="f", digits=2, big.mark=","))`
* Number of different sports funded: `r length(unique(sg_df$Sport))`
* Most funded sport: `r sp_vec[1,]` award (`r paste0(formatC(sp_per, format="f", digits=2),"%")`)
* Most funded organisation: `r sp_org[1,1:2]`
* Predominant focus was: `r sp_fo[1,1]` (`r sp_fo[1,2]`)

<br>

## Summary

```{r summary, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
table_df <- sg_df %>%
  rename("URN" = GMS_URN, `Award Amount` =  AWARD, `Total Project Cost` = Total_Project_Cost,  `Partnership Funding` = Partnership_Funding, `Other Funding` = OTHER_FUNDING, `Government Outcome` = Govt_Outcome, "Region" = GOVERNMENT_OFFICE_REGION, `Organisation Type`= Org_Class, `IMD Decile` = "IMD_Decile_2019", `Active Lives Survey Inactive (%)` = ALS_Inactive_PC_2018, `Total Population` = Population..2018.Estimates....Total.population.) %>%
  select(c("Award Amount", "Total Project Cost", "Sport", "Region", "Organisation Type", "Focus", "IMD Decile", "Total Population","Active Lives Survey Inactive (%)"))
kable(summary(table_df), row.names = 0, col.names = c("Award Amount", "Total Project Cost", "Sport", "Region", "Organisation Type", "Focus", "IMD Decile", "Total Population","Active Lives Survey Inactive (%)")) %>%
  kable_styling(bootstrap_options = "bordered",
                full_width = FALSE) 
```

## Award amounts
```{r award, echo=FALSE, fig.align="center", message=FALSE, warning=FALSE}
ggplot(sg_df)+
  geom_point(aes(AWARD, Total_Project_Cost), color = "#5D4086")+
  theme(text = element_text(family = "Norwester", size = 12, color = "#525252"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left")  +
  scale_color_manual(values = se_colour, name = "Focus")+
  labs( x= "Amount Awarded", y = "Total Project Cost", title = "Funding Amount by Total Project Cost")+
  scale_y_continuous(labels = scales::dollar_format(prefix = "£")) +
  scale_x_continuous(labels = scales::dollar_format(prefix = "£")) 
```

<br>

This graph shows the distribution of a the awards. It visualises five summary statistics (the median, two hinges and two whiskers), and all "outlying" points individually

```{r award_box, echo=FALSE, fig.align="center", message=FALSE, warning=FALSE}
ggplot(sg_df)+
  geom_boxplot(aes(x = "", y = AWARD) , color = "#5D4086", outlier.colour =  "#5D4086" )+
  theme(text = element_text(family = "Norwester", size = 12, color = "#525252"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left")  +
  scale_color_manual(values = se_colour, name = "Focus")+
  labs( x= "", y = "Amount Awarded", title = "Funding Amount")+
  scale_y_continuous(labels = scales::dollar_format(prefix = "£"))
```


## Times funded

This graph shows the number of organisations who had been funded by the Small Grants by the number of projects they have been funded for. As per the below, the maximum number of separate funded projects for a single organisation was `r max(sp_org$n)`

```{r fund, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}

sp_org$n <- as.factor(sp_org$n)
ggplot(sp_org)+
  geom_bar(aes(x="", y = n, fill = n), stat = "identity" ) +
  coord_polar("y", start = 0)+
  theme(text = element_text(family = "Norwester", size = 12, color = "#525252"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left") +
  labs( x= "", y = "", title = "Number of times an Organisation was Funded", caption = "Funded by Small Grants during current strategy period")+
  scale_fill_manual(values = se_colour, name = "Number of times")

```

## Focus
```{r fo_pie, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center'}
ggplot(sg_df)+
  geom_bar(aes(x="", y = AWARD, fill = Focus), stat = "identity" ) +
  coord_polar("y", start = 0)+
  theme(text = element_text(family = "Norwester", size = 12, color = "#525252"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left") +
  labs( x= "", y = "", title = "Amount Awarded per Focus")+
    scale_fill_manual(values = se_colour)
```

## Sport

The below chart shows the amount awarded to each sport (top `r count_of_sp` awarded sports):

```{r sport2, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center'}

## as dataframe
sg_sport <- as.data.frame(sg_sport)

sg_sport <- sg_sport[order(sg_sport$AWARD),] 
sg_sport %>%
  top_n(count_of_sp)%>%
  ggplot(aes(x= reorder(Sport, AWARD), y =AWARD))+
  geom_bar(stat = "identity", fill = "#5D4086") +
  coord_flip()+
  theme(text = element_text(family = "Norwester", size = 12, color = "#525252"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.text.y = element_text(hjust = 1, size = 8),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left") +
  labs( x= "Organisation type", y = "Number of Awards", title = "Type of Sport by Award Amount") +
  scale_y_continuous(labels = scales::dollar_format(prefix = "£"))
```

<br>

The below chart shows the amount awarded to each sport (bottom `r count_of_sp` awarded sports):

```{r sport3, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center'}

sg_sport <- sg_sport[order(sg_sport$AWARD, decreasing = TRUE),]
sg_sport %>%
  top_n(-count_of_sp)%>%
  ggplot(aes(x= reorder(Sport, AWARD), y =AWARD))+
  geom_bar(stat = "identity", fill = "#5D4086") +
  coord_flip()+
  theme(text = element_text(family = "Norwester", size = 12, color = "#525252"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.text.y = element_text(hjust = 1, size = 8),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left") +
  labs( x= "Organisation type", y = "Number of Awards", title = "Type of Sport by Award Amount") +
  scale_y_continuous(labels = scales::dollar_format(prefix = "£"))
```

## Organisation Type

```{r org_amount, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center'}

  ggplot(data = sg_org, aes(x= reorder(Org_Class, AWARD), y = AWARD))+
  geom_bar(stat = "identity", fill = "#5D4086") +
  coord_flip()+
  theme(text = element_text(family = "Norwester", size = 12, color = "#525252"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.text.y = element_text(hjust = 1, size = 8),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left") +
  labs( x= "Organisation type", y = "Award amount", title = "Organisation Type by Award Amount", caption = "Only one type has been selected per award. /nThis may not have been consistently applied by applicants")+
  scale_y_continuous(labels = scales::dollar_format(prefix = "£"))
  
```

## Government Outcome
The most invested in Government Outcome is `r gov_ord[1,1]` with `r gov_ord[1,2]` awards, and `r paste0("£", formatC(gov_ord[1,3], format="f", digits=2, big.mark=","))` total amount awarded.

```{r gov, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center'}
ggplot(gov)+
  geom_bar(aes(x=Govt_Outcome, y = AWARD, fill = Govt_Outcome), stat = "identity" ) +
  theme(text = element_text(family = "Norwester", size = 12, color = "#525252"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.text.y = element_text(hjust = 1, size = 8),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left") +
  labs( x= "Government Outcome", y = "Award Amount", title = "Total Award by Government Outcome") +
  scale_y_continuous(labels = scales::dollar_format(prefix = "£"))+
  scale_fill_manual(values = se_colour, name = "Outcome")
```
 

## Inactivity
This graph shows how much we invest into areas (LSOAs) by the proportion of those who are inactive (do less than 30 minutes of activity a week) in that area. The national rate of inactivity is 24.8%. The Active Live Survey report can be found [here](https://www.sportengland.org/media/14239/active-lives-adult-may-18-19-report.pdf)
```{r als, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
sg_df%>%
mutate(ALS_Inactive = case_when(ALS_Inactive_PC_2018 >= 0.9 ~ "6. Between 90% and 100%",
                                ALS_Inactive_PC_2018 >= 0.8 ~ "5. Between 80% and 89%",
                                ALS_Inactive_PC_2018 >= 0.7 ~ "4. Between 70% and 79%",
                                ALS_Inactive_PC_2018 >= 0.5 ~ "3. Between 50% and 69%",
                                ALS_Inactive_PC_2018 >= 0.3 ~ "2. Between 30% and 49%",
                              TRUE ~ '1. Under 30%'))%>%
  ggplot()+
  geom_bar(aes(x= ALS_Inactive, y = AWARD), fill = "#5D4086", stat = "identity" )+
  theme(text = element_text(family = "Norwester", size = 12, color = "#5F5F5F"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left")  +
  labs( x= "Percentage of Inactive (grouped)", y = "Amount awarded", title = "Active Lives Inactive",
        caption = ("Active Lives Survey 2018-2019")) +
  scale_y_continuous(labels = scales::dollar_format(prefix = "£")) 
```


## Data table
```{r dt, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
table_df <- sg_df %>%
  rename("URN" = GMS_URN, `Award Amount` =  AWARD, `Total Project Cost` = Total_Project_Cost,  `Partnership Funding` = Partnership_Funding, `Other Funding` = OTHER_FUNDING, `Government Outcome` = Govt_Outcome, "Region" = GOVERNMENT_OFFICE_REGION, `Organisation Type`= Org_Class, `IMD Decile` = "IMD_Decile_2019", `Active Lives Survey Inactive (%)` = ALS_Inactive_PC_2018, `Total Population (LSOA)` = Population..2018.Estimates....Total.population.) %>%
  select(c("URN",`Award Amount`, `Total Project Cost`, "Sport", "Region", `Organisation Type`, "Focus", `IMD Decile`, `Total Population (LSOA)`,`Active Lives Survey Inactive (%)`))

DT::datatable(table_df,  filter = 'top',
              options = list(pageLength = 5,
                                       lengthMenu = c(5, 10, 15, 20)))%>%
  formatCurrency(c(2,3), '£') %>%
  formatPercentage(10, 2) %>%
  formatRound(9,0) %>%
formatStyle(columns = "URN", font = "Norwester", color = "#B8345C" , backgroundColor = "#E79AB5" )
  

```

<br>

# Funding across the country {.tabset .tabset-fade .tabset-pills} 

Distribution of funding throughout England 

<style>
.leaflet {
    margin: auto;
}
</style>

## Individual projects

```{r map, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center'}
sg_map <- sg_df %>%
mutate(group = cut(AWARD, breaks = c(0, 500, 1000, 2000, 50000, 7500, 10000, Inf), 
                   labels = c("0 - £500", "£501 - £1,000", "£1,001 - £2,000", "£2,001 - £5,000", "£5,001 - £7,500", "£7,501 - £10,000", "Over £10,000"))) 

pal <- colorFactor(palette = c('#fff7f3','#fde0dd','#fcc5c0','#fa9fb5','#f768a1','#dd3497','#ae017e','#7a0177'), domain = sg_map$group)


leaflet() %>% addTiles() %>%
  addCircleMarkers(data= sg_map, ~longitude, ~latitude, 
                   popup = paste("Amount:", "£", formatC(as.numeric(sg_map$AWARD), big.mark = ",", digits = 10),  "<br>",
                                 "Name:", sg_map$ORGANISATION_NAME), 
                   color = ~pal(group),
                   fillOpacity = ~pal(group),
                   radius = ~sqrt(Total_Project_Cost)* 0.01) %>%
  addLegend("bottomright", pal = pal, values = sg_map$group, title = "Small Grants Awards: Sport England 2019",
            opacity = 1)
```

<br>

## Awards by Local Authority

The top LA for number of awards is: `r sp_la[1,1]` with `r sp_la[1,2]` awards.

```{r lamap, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.align="center"}
## For map
library(rgdal)
library(viridis)
library(plyr)

setwd("input/Local_Authority_Districts_April_2019_Boundaries_UK_BUC")
map1 <- readOGR(dsn = ".", layer = "Local_Authority_Districts_April_2019_Boundaries_UK_BUC", verbose = FALSE)

sg_df$lad19nm <- sg_df$LOCAL_AUTHORITY
df <- sg_df


## map without distortion
map1@data$id <- rownames(map1@data)
map1@data   <- join(map1@data, df, by="lad19nm")
map1$AWARD[is.na(map1$AWARD)] <- 0
map.clean     <- fortify(map1)
map.clean     <- join(map.clean,map1@data, by="id")

map.clean <- map.clean[grep("E", map.clean$lad19cd),] # England only
map.clean <- map.clean[, !duplicated(colnames(map.clean))] # remove dupe cols


p <- ggplot() +
  geom_polygon(data = map.clean, aes(x = long, y = lat, group = group, fill = AWARD))+
  theme_minimal() + 
  theme(axis.title.x = element_blank(),axis.title.y = element_blank(), 
        text = element_text(family = "Norwester", color = "#5F5F5F"), 
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5),
        panel.grid = element_blank(), 
        axis.text = element_blank(),
        axis.ticks.x = element_blank())+
  scale_fill_gradient(high = '#4a1486' , low = '#f2f0f7' ,
                      name='Award Amount', labels = scales::dollar_format(prefix = "£"))
p

```

<br>

# Environmental Indicators

We can also look at a number of environmental indicators that impact on the communities that the funding is targeting. The Index of Multiple Deprivation is the aggregate of all the different indicators (Crime, Employment, etc.), and this Index gives a relative score of deprivation in a locale.
The project in the most deprived area (`r sp_la_imd$LOCAL_AUTHORITY[1]`) is titled "`r sp_la_imd$PROJECT_TITLE[1]`", and was awarded `r paste0("£", formatC(mean(sp_la_imd$AWARD[1]), format = "f", digits=2, big.mark=","))`
The below chart shows the distribution of projects by IMD decile (2019).

```{r imd, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
ggplot(sg_df)+
  geom_histogram(aes(IMD_Decile_2019), fill = "#00818F", binwidth = 1 )+
  theme(text = element_text(family = "Norwester", size = 12, color = "#5F5F5F"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left")  +
  scale_color_manual(values = se_colour, name = "Government\n Outcome")+
  labs( x= "Decile", y = "Number of awards", title = "IMD Decile Distribution",
        caption = ("1 is most deprived, 10 is least deprived"))+
  scale_x_continuous(labels = scales::comma)+
  scale_y_continuous(labels = scales::comma)
  

```

<br>

# Measure and Evaluation: Successful projects {.tabset .tabset-fade .tabset-pills} 

What can we understand about a project to help us identify which projects have been successful  so that we can invest in projects that are more likely to be successful in the future?

Success here is defined as projects that have met or exceeded their 12 month targets. `r tab_me[2]` are successful out of the `r nrow(me_df)` projects had M&E data.

## Overall

The below table and graph shows the average target per project, and the average number of participants at 12 months.
```{r success_mean,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
kable(as.data.frame(colMeans(me_df_mean)), col.names = "Figures Overall") %>%
  kable_styling(bootstrap_options = "bordered",
                full_width = FALSE) 
```
<br>


```{r mean_df_1, success,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
melt(me_df_mean[,1:2]) %>%
ggplot(aes(x=variable, y =value))+
  geom_bar(aes( fill = variable),stat = "identity") +
  theme(text = element_text(family = "Norwester", size = 12, color = "#525252"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        panel.grid = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left") +
  labs( x= "Target / Actual Participants ", y = "Number of Participants", title = "Targets and actuals", caption = "'Successful' is defined as projects that \nmeet or exceed the year 1 target" )+
  scale_fill_manual(values = se_colour, name = "From a project that \nmet targets")+
  scale_y_continuous(labels = scales::comma_format(big.mark = ",")) 
```

<br>
This graph shows the proportion of awards that were "successful" or not at the 12 month point:
```{r success,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
me_df <- me_df %>%
  select("GMS_URN", "Sport", "Org_Class", "AWARD", "PROJECT_COST", "PARTICIPANTS BASELINE", "PARTICIPANTS PROJECT OVERALL" , "YEAR 1 PARTICIPANTS TARGET", "12 MONTH PARTICIPANTS", "prediction") %>%
  mutate(`Number of Participants (over / under target)` = `12 MONTH PARTICIPANTS` - `YEAR 1 PARTICIPANTS TARGET`)%>%
  mutate(`Success (Yes / No)` = if_else(`Number of Participants (over / under target)` >0, "Yes", "No"))
  
ggplot(me_df)+
  geom_bar(aes(x="", y =AWARD, fill = prediction), stat = "identity" ) +
  coord_polar("y", start = 0)+
  theme(text = element_text(family = "Norwester", size = 12, color = "#525252"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left") +
  labs( x= "", y = "", title = "Proportion of Amount Awarded to 'Successful' Projects", caption = "'Successful' is defined as projects that \nmeet or exceed the year 1 target" )+
  scale_fill_manual(values = se_colour, name = "Is the project \nsuccessful?")

```

<br>
The below boxplot shows the *distribution* of award amounts by whether the project was successful or not:
```{r success_box, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
ggplot(me_df)+
  geom_boxplot(aes(x="" , y = AWARD, fill = prediction) )+
  theme(text = element_text(family = "Norwester", size = 12, color = "#5F5F5F"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left")  +
  labs( x= "Successful or Not?", y = "Amount awarded", title = "Amount Awarded by Success of Project") +
  scale_y_continuous(labels = scales::dollar_format(prefix = "£")) +
  scale_fill_manual(values = se_colour, name = "Is the project \nsuccessful?")
```

## Gender

The below table shows the average target numbers and proportions by gender:

```{r women_target,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}

me_df_women.num <- me_df_women[,1:4]
kable(as.data.frame(colMeans(me_df_women.num)), col.names = "Figures Overall")%>%
  kable_styling(bootstrap_options = "bordered",
                full_width = FALSE) 

```

<br>

The below table shows the average 12 month actual numbers and proportions by gender:

```{r women_actuals,  echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}

me_df_women.num <- me_df_women[,5:8]

kable(as.data.frame(colMeans(me_df_women.num)), col.names = "Figures Overall")%>%
  kable_styling(bootstrap_options = "bordered",
                full_width = FALSE) 

```

<br>

The below bar chart shows the number or participants after 12 months, and shows the proportion of these that came from either 'successful' projects or those that had 'failed' i.e. did not meet their 12 month target:
```{r women_bar, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
melt_df_women <- melt(me_df_women[,c(5,6,9)])

ggplot(melt_df_women, aes(x=variable, y =value))+
  geom_bar(aes( fill = prediction),stat = "identity") +
  theme(text = element_text(family = "Norwester", size = 12, color = "#525252"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        panel.grid = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left") +
  labs( x= "Gender of Participants ", y = "Number of Participants", title = "Targets by Gender", caption = "'Successful' is defined as projects that \nmeet or exceed the year 1 target" )+
  scale_fill_manual(values = se_colour, name = "From a project that \nmet targets")+
  scale_y_continuous(labels = scales::comma_format(big.mark = ",")) 
```


## Sport
This barchart shows the amount awarded by the 5 most awarded sports, and shows the proportion that were "successful" (of those with M&E data):
```{r me_sport_bar, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
me_df %>%
  filter(Sport == c("Sailing & Yachting", "Rowing", "Association Football", "Exercise & Fitness", "Boxing" ))%>%
  ggplot()+
  geom_col(aes(x=Sport , y = AWARD, fill = prediction) )+
  theme(text = element_text(family = "Norwester", size = 12, color = "#5F5F5F"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left")  +
  labs( x= "Successful or Not?", y = "Amount awarded", title = "Award Amount by Sport") +
  scale_y_continuous(labels = scales::dollar_format(prefix = "£")) +
  scale_fill_manual(values = se_colour, name = "Is the project \nsuccessful?")
```
<br>

The below boxplot shows the distribution of award amounts by project success (of those with M&E data), and the 5 most awarded sports:
```{r me_sport, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
me_df %>%
  filter(Sport == c("Sailing & Yachting", "Rowing", "Association Football", "Exercise & Fitness", "Boxing" ))%>%
  ggplot()+
  geom_boxplot(aes(x=Sport , y = AWARD, fill = prediction) )+
  theme(text = element_text(family = "Norwester", size = 12, color = "#5F5F5F"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left")  +
  labs( x= "Successful or Not?", y = "Amount awarded", title = "Award Amount by Sport") +
  scale_y_continuous(labels = scales::dollar_format(prefix = "£")) +
  scale_fill_manual(values = se_colour, name = "Is the project \nsuccessful?")
```

## Organisation type
```{r me_org_bar, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
me_df %>%
  filter(Org_Class == c("Unregistered Club or Association", "Registered Charity", "Registered CASC", "Company Limited by Guarantee (without share capital)", "Charitable Incorporated Organisation" ))%>%
  ggplot()+
  geom_col(aes(x=Org_Class , y = AWARD, fill = prediction) )+
  theme(text = element_text(family = "Norwester", size = 12, color = "#5F5F5F"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "bottom")  +
  labs( x= "Successful or Not?", y = "Amount awarded", title = "Award Amount by Organisation Type", caption = "Only one type has been selected per award. \nThis may not have been consistently applied by applicants") +
  scale_y_continuous(labels = scales::dollar_format(prefix = "£")) +
  scale_fill_manual(values = se_colour, name = "Is the project \nsuccessful?")
```

<br>


```{r me_org_box, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}

me_df %>%
    filter(Org_Class == c("Unregistered Club or Association", "Registered Charity", "Registered CASC", "Company Limited by Guarantee (without share capital)", "Charitable Incorporated Organisation" ))%>%
  ggplot()+
  geom_boxplot(aes(x=Org_Class , y = AWARD, fill = prediction) )+
  theme(text = element_text(family = "Norwester", size = 12, color = "#5F5F5F"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "bottom")  +
  labs( x= "Successful or Not?", y = "Amount awarded", title = "Award Amount by Organisation Type", caption = "Only one type has been selected per award. \nThis may not have been consistently applied by applicants") +
  scale_y_continuous(labels = scales::dollar_format(prefix = "£")) +
  scale_fill_manual(values = se_colour, name = "Is the project \nsuccessful?")

```

## Data table
```{r me_dt, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
me_tab <- me_df %>%
  select("GMS_URN", "Sport", "Org_Class", "AWARD", "PARTICIPANTS BASELINE", "YEAR 1 PARTICIPANTS TARGET", "12 MONTH PARTICIPANTS", "prediction") %>%
  mutate(`Number of Participants (over / under target)` = `12 MONTH PARTICIPANTS` - `YEAR 1 PARTICIPANTS TARGET`)%>%
  mutate("Percentage (over / under target" = `Number of Participants (over / under target)` / `YEAR 1 PARTICIPANTS TARGET`)

DT::datatable(me_tab, colnames = c('Row', 'URN', 'Sport', 'Organisation Type', 'Award Amount', 'Baseline', 'Year 1 Target', 'Year 1 Participants', 'Outcome', 'Number: over/under target'),  filter = 'top',
              options = list(pageLength = 5,
                                       lengthMenu = c(5, 10, 15, 20)))%>%
  formatCurrency(4, '£') %>%
  formatRound(c(5:8, 10),0) %>%
formatStyle(columns = "GMS_URN", font = "Norwester", color = "#B8345C" , backgroundColor = "#E79AB5" )

```

<br>

# Text Analysis 
Using the "12 month progress" field, we can look at the unstructured data of the Measure and Evaluation element of an award. We can do this using word frequency, topic extraction using LDA (Latent Dirichlet Modelling), and also assigning sentiment. 

<br>

## Term frequency score
The wordcloud was produced through text processing ) which stripped out punctuation and non-relevant words, and the word frequency was weighted with the 'importance' given to each word.

```{r 12_prog, echo=FALSE, fig.align="center", message=FALSE, warning=FALSE}

wordcloud2(text_df[,c(2,4)], fontFamily = "Norwester", color = "random-light")

```

<br>

## Topics

The topic modelling was created using the LDA model, and the "Elbow" method applied to assign the number of topics (more information at this [site](https://www.knime.com/blog/topic-extraction-optimizing-the-number-of-topics-with-the-elbow-method)).

```{r 12_prog_top, echo=FALSE, fig.align="center", message=FALSE, warning=FALSE}

kable(topic_tbl) %>%
  kable_styling(bootstrap_options = "bordered",
                full_width = FALSE) 
```

<br>

## Sentiment
Sentiment score using a lexicon-based approach: The approach assigns a sentiment to each word of positive or negative. This can be translated int a score (number of positive words - number of negative words) / total number of words).

The number of Small Grants projects with a comment in the "12 month progress" field: `r nrow(sentiment_df)`

<br>

### Proportion of projects with positive / negative sentiment

* Positive comments: `r sent_tab[1]`
* Negative comments: `r sent_tab[2]`

```{r Sent_pie, echo=FALSE, fig.align="center", message=FALSE, warning=FALSE}
ggplot(sentiment_df)+
  geom_bar(aes(x="", y = Sentiment, fill = Sentiment), stat = "identity" ) +
  coord_polar("y", start = 0)+
  theme(text = element_text(family = "Norwester", size = 12, color = "#525252"),
        plot.title=element_text(size=14,family = "Norwester", face="bold", hjust = 0.5, color = "#525252"),
        axis.text.x = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.caption=element_text(size=9, hjust=0, margin=margin(15,0,0,0)),
        legend.position = "left") +
  labs( x= "", y = "", title = "Sentiment")+
    scale_fill_manual(values = se_colour)
```

<br>

### Example of sentiment scoring

```{r Sent, echo=FALSE, fig.align="center", message=FALSE, warning=FALSE}

kable(sentiment_df[1,]) %>%
  kable_styling(bootstrap_options = "bordered",
                full_width = FALSE) 

```
